### 2024-12-18 skripsi/chapter4.tex
modify ch1 aims, add ch4 implementation & scenarios, start ch4 experimental results

The computational model is divided into four large phases. The processes within these phases are based on the broad outcomes of the phase. The phases themselves are data preparation, data transformation, model training, and model evaluation. Each phase is composed of processes that together works toward the output of the larger process.

@@

\subsection{Design}
\noindent In this subsection, the architectural design of the software product is explained. The architecture consists of modules with similar concerns. This was chosen because of the intuition built by the use of other libraries with a similar architecture. The broad aim of modeling PDEs with LSSVR and basis functions coefficients can be separated into smaller concerns of data, model, basis functions, and utilities.

@@

\subsection{Testing}

\section{Experimental Scenarios}
\noindent

\section{Experimental Results}
\noindent

% \section{Sub Bab Hasil}
% \noindent Bab ini memaparkan pekerjaan penelitian dan, terutama, hasil-hasilnya, untuk dianalisis. Secara komprehensif bab ini merepre-sentasikan curahan pemikiran dan kemampuan mahasiswa dalam menjalani pekerjaan penelitian, yang hasil-hasilnya dapat dipertanggungjawabkan. Banyak pendukung yang diperlukan dalam penulisan bab ini, seperti skema penting pengolahan data, penurunan model matematika, asumsi khusus, tabulasi hasil dan analisis, dan gambar atau grafik yang membantu dalam paparan analisis. Judul bab dan sub bab disesuaikan dengan isi paparan.

% \section{Memasukkan Gambar}
% \noindent Setiap gambar harus dirujuk pada naskah TA, termasuk gambar pada Lam- piran, menggunakan huruf pertama kapital (G) dan nomor gambar, tidak berdasarkan posisi relatifnya (misalnya di bawah ini atau sebelum ini). Format gambar yang umum adalah jpg, png, dan postscript (ps atau eps). Ukuran huruf pada nama sumbu dan label figures/grafik harus cukup besar dan jelas, demikian halnya dengan angka pada sumbu. Gambar dan grafik dapat berwarna dengan pilihan warna yang tegas dan jelas.

% \subsection{Contoh Gambar Sederhana}
% Contoh menginput gambar pada buku TA pada
% \begin{figure}[ht] %h artinya here!
%     \centering
%     \includegraphics[width=0.3\linewidth]{figures/Semiconductor-Fermi-Level-Band-Diagram-1.jpg}
%     \caption{Tingkatan Fermi pada\\Bahan Semikonduktor}\label{fermilevel}
% \end{figure}
% Apabila ada dua gambar, kita juga bisa menaruh keduanya berdampingan.

% \begin{figure}[H]
%     \centering
%     \subfigure[]{
%         \includegraphics[width=0.4\linewidth]{figures/Semiconductor-Fermi-Level-Band-Diagram-1.jpg}\label{surf-1}
%     }\hspace{0.1\linewidth}
%     \subfigure[]{
%         \includegraphics[width=0.4\linewidth]{figures/Semiconductor-Fermi-Level-Band-Diagram-1.jpg}\label{surf-2}
%     }
%     \caption{Dengan menempatkan gambar \subref{surf-1} dan \subref{surf-2}, pembaca akan lebih mudah membandingkan keduanya.}\label{surface}
% \end{figure}

>>>>

The computational model is divided into four large phases. The processes within these phases are based on the broad outcomes of the phase. The phases themselves are data preparation, data transformation, model training, and model evaluation. Each phase is composed of processes that together works toward the output of the larger process. At the end, a final process of analysis and interpretation is carried out.

@@

\subsection{Design} \label{sec:implementation_design}
\noindent In this subsection, the architectural design of the software product is explained. The architecture consists of modules with similar concerns. This was chosen because of the intuition built by the use of other libraries with a similar architecture. The broad aim of modeling PDEs with LSSVR and basis functions coefficients can be separated into smaller concerns of data, model, and basis functions. Each of these concerns are addressed by a corresponding module. The explanation of each module are as follows:
\begin{enumerate}
  \item Utilities

    \noindent This module provides basic tools that are used across the entire library. The utilities this module needs to provide include scaling, conversion between real and complex representations of matrices, resizing coefficient tensors, and adaptors for or the traditional ordinary differential equation solvers. Scaling functionality is used for inputs into the support vector machines. And as previously mentioned, since LSSVR only works with real numbers, the tools for converting between complex and their real number representations is also necessary. Another useful tool for working with basis coefficients is a way to resize the coefficients in terms of how many basis functions is desired. An example of the usage is downsizing Fourier coefficients from 12 to 8 modes. The last utility mentioned here provides the ability to solve ordinary differential equations. This utility is useful for generating data for certain types of PDEs such as the Burgers' equation.

  \item Basis Functions

    \noindent As the proposed model will learn the relationships defined by PDEs within the spectral domain, a large proportion of the computation will also concern the use of the basis functions themselves. The module will define the functionality and data structure that is expected to be used when working with basis functions in this library. First and foremost, the central function of this module is to store and facilitate the connection between the coefficients in the spectral domain and their values in the physical domain. This is dependent on the basis functions used. For the Fourier basis that we use in this study, this means that the transformations are the discrete Fourier transform and the discrete inverse Fourier transform. Other than this core functionality, information on the characteristics of the coefficients such as the number of modes, in other words basis functions, should also be provided by this module. In addition, other functionality included in this module are visualization of coefficients in physical space, generation of random functions, time dependent \& complex/real coefficients, noise perturbation of functions, and differential operators.

  \item Model

    \noindent The proposed SpectralSVR model is implemented in this module. As the proposed model essentially wraps a support vector regression model in such a way that it can process coefficients, this module will house the SpectralSVR wrapper itself and the LSSVR variant of support vector regression that is used in this study. The SpectralSVR wrapper must provide three core pieces of functionality. First, it must translate complex input and output samples into their real representations. This applies to both training and evaluation. Second, in addition to predicting the coefficients during evaluation, the wrapper should also be able to predict the function values themselves for any given set of point coordinates. This is important for quick evaluations of a predicted function at an arbitrary point. The third functionality is inverse input function search for a given output. This is useful for inverse problems where the parameters are unknown, but the solution is partially known. Aside from the core functionality, a complementary functionality is testing for the learned model.

    The model module also implements the LSSVR model. This implementation provides two main functionalities. The first is training the LSSVR model with some given model hyperparameters such as the regularization constant and kernel. The second functionality is the evaluation of the learned model. Both functionalities allow for multiple output regression. This is required to learn the mappings for many coefficients at once. Other functionality provided by the implementation include interpretability functions and model serialization for saving and loading the learned model to disk.

  \item Problems

    \noindent This module provides the functionality concerned with generating data. Two out of the three datasets in this study are generated with this module. In general this module provides two pieces of functionality. The first is generation of the data itself. This functionality generates at least two sets of functions that are related to each other by the chosen PDE\@. The generation is based on at least two parameters which are the number of samples to generate and the number of modes or basis functions to generate the coefficients for. The second functionality provided by this module is the residual based on the chosen PDE\@. This is useful to verify if any set of functions comply with the PDE\@. For example, the residual may be used to verify the generated data. Another use is to grade the predictions by the learned model.

The design presented here is the product of multiple iterations that started with building the core model and then the basis functions and finally the data. Throughout the iterations, the utilities were also added to alleviate the need for general tools.

@@

\noindent The design in \lccref{sec:implementation_design} is implemented within the same iterative loop. As such, the feedback from tests is immediate, and any fixes can be implemented quickly. The implementation is done using the Python language and a number of Python libraries. To facilitate the management of dependencies, the implementation makes use of the package management tool Poetry \autocite{eustacePoetryPythonPackaging2024}. In addition, the Python version is managed using the tool pyenv. The implementation process starts with installation of the required dependencies. The dependencies are specified in the \verb|pyproject.toml| as shown in \lccref{fig:poetry_add_deps}.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Toml]
...
[tool.poetry.dependencies]
python = "^3.10.6"
numpy = "^1.24.0"
tqdm = "^4.65.0"
jaxtyping = "^0.2.20"
pandas = "^2.2.2"
torchmetrics = "^1.4.2"
scikit-learn = "^1.5.0"
matplotlib = "^3.9.0"
torch = {version = "^2.4.1+cu124", source = "pytorch-gpu"}
torchvision = {version = "^0.19.1+cu124", source = "pytorch-gpu"}
torchaudio = {version = "^2.4.1+cu124", source = "pytorch-gpu"}
ray = {extras = ["tune"], version = "^2.37.0"}
hyperopt = "^0.2.7"
ipywidgets = "^8.1.5"
torchdiffeq = "^0.2.4"
xarray = "^2024.10.0"
zarr = "^2.18.3"
gcsfs = "^2024.10.0"

[[tool.poetry.source]]
name = "pytorch-gpu"
url = "https://download.pytorch.org/whl/cu124"
priority = "explicit"

[build-system]
requires = ["poetry-core"]
build-backend = "poetry.core.masonry.api"
  \end{lstlisting}
  \caption{Contents of pyproject.toml configuration file that define the dependencies using Poetry}\label{fig:poetry_add_deps}
\end{figure}

After ensuring that the correct version of python is being used with pyenv, the dependencies specified in the \verb|pyproject.toml| file can be installed using the shell command \verb|poetry install|. At the end of each iteration, the dependencies are subject to change depending on whether any new ones are needed or old dependencies can be removed.

In the following, we will discuss the implementation of each module and how the phases of computational model relates to each implemented functionality. The base data structure that is used in the implementation is the Pytorch Tensor \autocite{Ansel_PyTorch_2_Faster_2024}. Tensors are a generalization of scalars and array-like structures such as vectors and matrices. This is essential since many computations that is done in the proposed model are based on matrices and higher dimensional arrays. Pytorch Tensors allow the elements to be of different data types such as floating point numbers of different precision or even complex numbers. In addition to basic operations for working with the Tensors such as matrix multiplications, Pytorch also provides many algorithms for use when working with Tensors such as linear equation solvers or the Fast Fourier Transform. Another reason for choosing to use Pytorch is the ability to parallelize computation using specialized processors such as GPUs. Lastly, Pytorch is widely used in the machine learning community which reflects on its reliability, wealth of community knowledge and support, and active development.

\subsubsection{Utilities}

\noindent The utilities module provide common tools that are used when working with the developed software. First, because the basis function used in this study is the Fourier basis, a large proportion of concern is the storage of Fourier basis coefficients. The Fourier series representation that is used in this study is the complex representation as show in \lccref{eq:fourier_series}. However, as previously mentioned, least squares support vector machines is constructed to work with real numbers. In order for LSSVR to learn the relationship between complex numbers, a representation of complex numbers in terms of real numbers is necessary. Since sample features and labels for LSSVR need to be flattened into one dimension which result in a matrix with rows of samples, the representation of complex numbers using real numbers can be done after features and labels have been flattened. An example of such a matrix of two samples with four features each can be seen in \lccref{eq:complex_matrix}. To represent the complex number features as real numbers, once can simply split a single complex number into two real numbers. When this is done to the entire matrix of complex numbers, the result is \lccref{eq:complex_matrix_real_rep}.
\begin{align}
  \begin{bmatrix}
    -1.03+0i & -0.52+0.39i & -0.54+0i & -0.52-0.39i\\
    -0.85+0i & -0.39+0.19i &  0.87+0i & -0.39-0.19i
  \end{bmatrix}\label{eq:complex_matrix}\\
  \begin{bmatrix}
    -1.03 & 0 & -0.52 & 0.39 & -0.54 & 0 & -0.52 & -0.39\\
    -0.85 & 0 & -0.39 & 0.19 &  0.87 & 0 & -0.39 & -0.19
  \end{bmatrix}\label{eq:complex_matrix_real_rep}
\end{align}
There are several ways to do this in Pytorch, such as extracting the real and imaginary element components into separate matrices and then interleaving the columns to create the real representation. However, a more performant version can be made by using built-in functions of the library that directly allows views of complex valued tensors as real tensors and vice versa. The \verb|view_as_real| and \verb|view_as_complex| functions allow the conversion between real and complex representations by representing complex valued tensors by adding a dimension of size two for each component of complex numbers. The resulting implementation can be seen in \lccref{fig:complex_matrix_conversion_implementation}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
def to_real_coeff(x: torch.Tensor) -> torch.Tensor:
    return torch.view_as_real(x).flatten(-2)

def to_complex_coeff(x: torch.Tensor) -> torch.Tensor:
  return torch.view_as_complex(
    x.reshape((*x.shape[:-1], -1, 2))
  )
  \end{lstlisting}
  \caption{Utility function to convert between complex matrices and the real representations}\label{fig:complex_matrix_conversion_implementation}
\end{figure}

Another important utility used on inputs for the proposed SpectralSVR is a scaler. This utility has been implemented in a variety of ways and using different tools. However, as the standard scaling is not provided by Pytorch, we have also implemented our own version. The scaler accepts Pytorch tensors or tuples of tensors of real or complex elements. First, the scaler is implemented as a class that is instantiated. The scaler is then fitted onto the reference data to obtain the standard deviation and mean of each feature. This information is stored as properties of the class. Any complex tensors are converted into their real representations using the functions in \lccref{fig:complex_matrix_conversion_implementation}. Once the scaler instance is fitted, any set of tensors with the same number of features (columns) and element type (complex or real) can be roughly transformed into a standard deviation of one and mean of zero for every feature. This is done by subtracting the fitted mean of each feature from the elements in the column and then dividing with the feature's standard deviation. The implementation also provides a method to retrieve a scaler fitted on a subset of the original tuple of tensors. Serialization for saving to and loading from disk is also implemented. This is done simply by using the Pytorch save and load functions on the scaler instance itself. For some use cases, an inverse scaling operation to the original standard deviation and mean is also essential. The implementation is simply the reverse of the transformation process where the elements are multiplied by the standard deviation and the mean is the added.

Another group of utility in the module is concerned with managing the number of modes a tensor of coefficients has. First, a function called \verb|resize_modes| accepts a tensor of coefficients and the target number of modes in each dimension. The first dimension is always assumed to represent the different samples. Because of this the target modes will only affect the dimensions after the first. The wave numbers are also assumed to be symmetric about zero such that the highest absolute values of the wave number is in the middle of each dimension. The wave number decreases symmetrically as you move towards the edges of the dimension and reaching the lowest absolute wave number at the edges. This is done to ensure compatibility with the way existing libraries such as Pytorch works with tensors of Fourier coefficients. Because of this, shrinking a tensor of coefficients along a dimension means removing the center of the tensor along that dimension. And expanding to a target number of modes that is larger means that the tensor is padded with zeros in the center along the desired dimension. In addition, the remaining coefficients may be rescaled in order to balance the effect of adding or removing coefficients. The implementation is presented in \lccrefs{fig:coefficient_shrink_impl,fig:coefficient_expand_impl}.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
for dim, (target_mode, current_mode) in enumerate(
    zip(target_modes, current_modes), 1
):
    device = x.device
    start_range = torch.tensor(
        range((target_mode - 1) // 2 + 1), dtype=torch.int
    ).to(device=device)
    end_range = torch.tensor(
        range(current_mode - target_mode // 2,
          current_mode
        ),
        dtype=torch.int,
    ).to(device=device)

    x_resized = torch.concat(
        (
            x_resized.index_select(dim, start_range),
            x_resized.index_select(dim, end_range),
        ),
        dim,
    )
  \end{lstlisting}
  \caption{Example of shrinking a tensor of samples' coefficients to a target number of modes}\label{fig:coefficient_shrink_impl}
\end{figure}
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
for dim, (target_mode, current_mode) in enumerate(
    zip(target_modes, current_modes), 1
):
    device = x.device
    start_range = torch.tensor(
        range((current_mode - 1) // 2 + 1), dtype=torch.int
    ).to(device=device)
    # make sure that end range is empty if
    # the coefficient is only size 1
    end_range = torch.tensor(
        range(current_mode // 2, current_mode)
        if current_mode > 1
        else range(0),
        dtype=torch.int,
    ).to(device=device)
    padding_size = target_mode - current_mode
    modes = list(x_resized.shape)
    modes[dim] = padding_size
    padding = torch.zeros(modes).to(x_resized)

    x_resized = torch.concat(
        (
            x_resized.index_select(dim, start_range),
            padding,
            x_resized.index_select(dim, end_range),
        ),
        dim,
    )
  \end{lstlisting}
  \caption{Example of expanding a tensor of samples' coefficients to a target number of modes}\label{fig:coefficient_expand_impl}
\end{figure}
Next, the second way to manage the size of a tensor along a dimension is to interpolate the current values. The implementation uses simple linear interpolation. This is especially useful for situations such as coefficients that are themselves functions of time. This is a common application such as when solving the initial condition problem of a function in space and time. One might represent the function as a Fourier series with basis as functions of space and coefficients as functions of time. In this case, the coefficients for time values that are not available can be interpolated. The implementation of this is shown in \lccref{fig:coefficient_interpolation_impl}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
    index_floor = index_float.floor().to(torch.int)
    index_ceil = index_float.ceil().to(torch.int)
    x_ceil = x.index_select(dim, index_ceil)
    x_floor = x.index_select(dim, index_floor)
    # interpolate coefficients
    index_shape = [1 for _ in range(x_floor.ndim)]
    index_shape[1] = -1
    index_scaler = (
        ((index_float - index_floor) / (
          index_ceil - index_floor
        ))
        .reshape(index_shape)
        .nan_to_num()
    )
    # ynt + scaler * (ynt1 - ynt)
    # (1 - scaler) * ynt + scaler * ynt1
    x_interp = torch.lerp(x_floor, x_ceil, index_scaler.to(x))
  \end{lstlisting}
  \caption{Example of interpolating for indices (floating point indices) between the available ones (intiger indices)}\label{fig:coefficient_interpolation_impl}
\end{figure}

Finally, this module also provides the types and reexports implementations of ordinary differential equation solvers from the torchdiffeq library \autocite{Chen_torchdiffeq_2021}. The implementation can be seen in \lccref{fig:solver_impl}.
\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
RHSFuncType = Callable[[torch.Tensor, torch.Tensor],
  torch.Tensor
]
SolverSignatureType = Callable[[
        RHSFuncType, torch.Tensor, torch.Tensor,
    ],
    torch.Tensor
]
MixedRHSFuncType = Callable[[
        torch.Tensor, torch.Tensor, torch.Tensor,
    ],
    torch.Tensor,
]
MixedSolverSignatureType = Callable[
    [MixedRHSFuncType, torch.Tensor, torch.Tensor],
    torch.Tensor,
]

def euler_solver(
    rhs_func: RHSFuncType,
    y0: torch.Tensor,
    t: torch.Tensor,
):
    solution = torch.zeros((len(t), *y0.shape)).to(y0)

    j = 1
    solution[j - 1] = y0
    for t0, t1 in zip(t[:-1], t[1:]):
        dt = t1 - t0
        y = solution[j - 1]
        solution[j] = y + dt * rhs_func(t0, y)
        assert (
            solution.isnan().sum() == 0
        ), f"solver encountered nan at timestep {j} (t={t0})"
        j = j + 1
    return solution

implicit_adams_solver: SolverSignatureType = partial(
    odeint, method="implicit_adams", options={"max_iters": 4}
)  # type: ignore

lsoda_solver: SolverSignatureType = partial(
    odeint, method="scipy_solver", options={"solver": "LSODA"}
)  # type: ignore
  \end{lstlisting}
  \caption{Implementation of ODE solvers and the types}\label{fig:solver_impl}
\end{figure}

\subsubsection{Basis Functions}

\noindent This module implements two classes. The first is the base Basis class. The second is a subclass implementing the Fourier basis specifically. These classes provide the functionality needed to have an easier time when working with basis functions and their coefficients. First, the core function of these classes is to provide a way to transition between the spectral domain and the physical domain. This is implemented in the base class by storing the coefficients of sample functions as a class property. The coefficients are assumed to be at least of one function.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    @staticmethod
    def transform(
        f: torch.Tensor,
        res: TransformResType | None = None,
        periodic: bool = True,
        periods: PeriodsInputType = None,
        allow_fft: bool = True,
    ) -> torch.Tensor:
        if not torch.is_complex(f):
            f = f * (1 + 0j)
        res = transformResType_to_tuple(res, tuple(f.shape[1:]))
        periods = periodsInputType_to_tuple(periods, f.shape[1:])
        # perform 1d transform over every dimension
        F = f
        for cdim in range(1, ndims):
            F = FourierBasis._ndim_transform(
                F,
                dim=cdim,
                func="forward",
                res=res[cdim - 1],
                periodic=periodic,
                period=periods[cdim - 1],
                allow_fft=allow_fft,
            )

        return F
...
  \end{lstlisting}
  \caption{Implementation of forward Fourier transform}\label{fig:fourier_transform_impl}
\end{figure}

The transitions between the coefficients and functions values are implemented as a function to transform function values to coefficients and an inverse transform function to compute function values from coefficients. The base class enforces this by defining abstract methods which any subclass will need to implement. The FourierBasis subclass then implements the transform and inverse transform functions specific for transitioning between function values and Fourier basis coefficients. As before, the implementation assumes that the first dimension indexes the samples and the rest represent the coefficient wave numbers. This means that the implementation must allow for multidimensional functions. For the Fourier transform and inverse transform, the multidimensional version is relatively simple to implement. The multidimensional Fourier transform essentially performs the one dimensional Fourier transform along one dimension. And then, using the result to perform the one dimensional Fourier transform along the next dimension. This process is repeated until the last dimension. The implementation of the forward transform is presented in \lccref{fig:fourier_transform_impl}. The inverse transform simply changes the \verb|func| parameter to \enquote{inverse}.

The Fourier transform along each dimension is done by flattening every other dimension into the sample dimension. This essentially means that the function values are \enquote{sliced} along the dimension currently being transformed, and that each \enquote{slice} is its own \enquote{sample}. Mathematically, this is simply due to the matrix multiplication of basis functions along one dimension with every other dimension. The implementation of this is shown in \lccref{fig:ndim_fourier_transform_impl}.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    @staticmethod
    def _ndim_transform(
        f: torch.Tensor,
        dim: int,
        func: Literal["forward", "inverse"],
        res: slice,
        periodic: bool,
        period: float,
        allow_fft: bool,
    ) -> torch.Tensor:
        # flatten so that each extra dimension is
        # treated as a separate "sample"
        # move dimension to transform to the end
        # so that it can stay intact after f is flatened
        f_transposed = f.moveaxis(dim, -1)
        # flatten so that the last dimension is intact
        f_flatened = f_transposed.flatten(0, -2)

        F_flattened = FourierBasis._raw_transform(
            f_flatened,
            func=func,
            res=res,
            periodic=periodic,
            period=period,
            allow_fft=allow_fft,
        )
        # unflatten so that the correct shape is returned
        F_transposed = F_flattened.reshape(
          (*f_transposed.shape[:-1], res.step)
        )
        F = F_transposed.moveaxis(-1, dim)

        return F
...
  \end{lstlisting}
  \caption{Implementation of n-dimensional Fourier transform for a specific dimension}\label{fig:ndim_fourier_transform_impl}
\end{figure}

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    @staticmethod
    def _raw_transform(
        f: torch.Tensor,
        func: Literal["forward", "inverse"],
        res: slice,
        periodic: bool,
        period: float,
        allow_fft: bool,
    ) -> torch.Tensor:
        match func:
            case "forward":
                sign = -1
            case "inverse":
                sign = 1
        mode = f.shape[1]
        domain_starts_at_0 = res.start == 0
        domain_end_equal_to_period = res.stop == period
        can_use_fft = (
            domain_starts_at_0
            and domain_end_equal_to_period
            and periodic
            and allow_fft
        )
        if can_use_fft:
            if func == "forward":
                F = torch.fft.fft(
                  f, dim=1, n=res.step, norm="backward"
                )
            elif func == "inverse":
                F = torch.fft.ifft(
                  f, dim=1, n=res.step, norm="forward"
                )
        else:
            if periodic:
                n = res.start + torch.arange(res.step).to(f)
                n = n / res.step * period
            else:
                n = torch.linspace(
                    res.start, res.stop, res.step
                ).to(f)
            e = FourierBasis.fn(
                n.view(-1, 1),
                mode,
                periods=period,
                constant=sign * 2j * torch.pi,
            )

            F = torch.mm(f, e.T)
        return F
...
  \end{lstlisting}
  \caption{Implementation of the one dimensional Fourier Transform}\label{fig:1d_fourier_transform_impl}
\end{figure}

The one dimensional Fourier transform itself is implemented as a raw transform function. This function accepts input of a matrix of values to transform, a parameter to indicate the type of transformation (inverse transform or forward transform), the evaluation boundary and number of grid points, whether the evaluation should assume the function is periodic, the size of the function domain, and if the Fast Fourier Transform is allowed to be used. The evaluation boundaries, actual function domain size, whether the function can be assumed to be periodic, and if FFTs is allowed to be used together determine the Discrete Fourier Transform algorithm that is used. The FFT algorithm is used if the folowing are all true: the evaluation starts at zero and ends with the same value as the period, the function can be assumed to be periodic, and FFT is allowed. Otherwise, the transform is computed naively using \lccrefs{eq:discrete_fourier_transform,eq:inverse_discrete_fourier_transform}. The naive approach allows for evaluating function values of coefficients at various grid sizes and evaluation boundaries. This is useful for being able to evaluate at resolutions other than that of the original discretized function or coefficients. One use for this is plotting the function at different parts or resolutions. The relationship with the physical domain means that the boundaries of the domain is important. This is because the function values that will be used may only be valid within certain boundaries. Because of this, the information of physical domain bounds must also be stored. This is done by storing the span of the function in each dimension. The implementation we have gone with doesn't store any other information than the span in each dimension. Because of this, the information would need to be stored outside the Basis class instance and any outputs or operations with the classes will need to take this into account. The implementation of the one dimensional Fourier transform is presented in \lccref{fig:1d_fourier_transform_impl}.

The secondary set of functionality provided by the basis classes are mathematical operators on the functions the coefficients represent. There are four operators implemented, which are the addition, subtraction, derivative, and antiderivative operators. These four operations can be further categorized into arithmetic and calculus operations. The arithmetic operations, which are addition and subtraction, are implemented by following how the actual mathematical operations would be carried out on two functions which are Fourier series as shown in \lccref{eq:fourier_series_addition}. For subtraction, the plus sign is simply replaced with the minues sign.
\begin{align}
  \sum_{k} \hat{u}_k e^{2\pi ikx} + \sum_{k} \hat{f}_k e^{2\pi ikx} =
  \sum_{k} \left(\hat{u}_k + \hat{f}_k\right) e^{2\pi ikx} \label{eq:fourier_series_addition}
\end{align}
The implementation leverages the operator overloading in Python classes. For addition, the overload should be done to the \verb|__add__| function. This function adds the current instance to another object which is the other basis instance for our use case. We also want the operation to return an independent instance without any references to the previous instances. To do this, all properties are coppied into a new instance. This is then used to perform the arithmetic operation. The implementation in \lccref{fig:basis_addition_impl} demonstrates the addition operation. For subtraction, a similar implementation is done using the \verb|__sub__| function and switching the plus sign with the minus sign.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def __add__(self, other: Self):
        if isinstance(other, self.__class__):
            if other.coeff is None:
                return self.copy()
            elif self.coeff is None:
                return other.copy()
            else:
                result = self.resize_modes(other)
                result.coeff = result.coeff + other.coeff
                return result
        else:
            raise TypeError(
                f"unsupported operand type(s) for +: '{
                    self.__class__
                }' and '{type(other)}'"
            )
...
  \end{lstlisting}
  \caption{Implementation of basis addition function}\label{fig:basis_addition_impl}
\end{figure}

The calculus operations for Fourier basis take advantage of the properties of the Fourier series. As discussed in \lccref{sec:fourier_discussion}, derivatives and integration become multiplication and division in the spectral space of Fourier series. Our implementation takes advantage of this as shown in \lccref{fig:basis_grad_impl}. Similar to how subtraction is just a modification of the addition operation, the integral or antiderivative operator is also implemented simply by changing the operation with the term multiplier from multiplication to division. % TODO: change the reference to specific equations for calculus in spectral domain

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def grad(self, dim: int = 0, ord: int = 1) -> Self:
        copy = self.copy()
        if dim == 0 and self.time_dependent:
            # time dependent use finite differences
            dt = self.periods[0] / (self.time_size - 1)
            coeff = copy.coeff
            for o in range(ord):
                coeff = torch.gradient(coeff, spacing=dt, dim=1)[0]
            copy.coeff = coeff
        else:
            if self.time_dependent:
                # disregard time dimension
                dim = dim - 1
            k = copy.wave_number(copy.modes[dim])
            multiplier_dims = [1 for _ in range(copy.ndim)]
            multiplier_dims[dim] = copy.modes[dim]
            if self.time_dependent:
                multiplier_dims = (1, *multiplier_dims)
            multiplier = (
                2
                * torch.pi
                * 1j
                * k.reshape(multiplier_dims).to(copy.coeff)
                / self.periods[dim]
            )
            multiplier = multiplier.pow(ord)
            coeff = copy.coeff.mul(multiplier)
            coeff[:, ..., 0] = torch.tensor(0 + 0j)
            copy.coeff = coeff
        return copy
...
  \end{lstlisting}
  \caption{Implementation of basis grad function}\label{fig:basis_grad_impl}
\end{figure}

The last group of functionality implemented in the Basis and FourierBasis classes provides convenience and often used procedures when working with basis functions. First, random coefficient generation for the Fourier basis is implemented. This is done as discussed in step 3 of the data generation process discussed in \lccref{sec:data_generation}. The implemented function shown in \lccref{fig:generate_coeff_impl} is then wrapped in another function that creates a new FourierBasis instance with the generated coefficients.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    @classmethod
    def generate_coeff(
        cls,
        n: int,
        modes: int | tuple[int, ...],
        generator: torch.Generator | None = None,
        random_func: Callable[..., torch.Tensor] = torch.randn,
        complex_funcs: bool = False,
        scale: bool = True,
    ) -> torch.Tensor:
        if isinstance(modes, int):
            modes = (modes,)
        n_modes = len(modes)
        assert n_modes > 0, "modes should have at least one element"
        random_func = partial(random_func, generator=generator)
        if complex_funcs:
            coeff = random_func((n, *modes), dtype=torch.complex64)
        else:
            coeff = random_func((n, *modes), dtype=torch.complex64)

            vals = cls.inv_transform(coeff)
            coeff = cls.transform(vals.real + 0j)
        if scale:
            scaler = torch.tensor(modes).sum() * 0.2
            coeff = coeff.mul(scaler)
        return coeff
...
  \end{lstlisting}
  \caption{Implementation of Fourier Basis coefficient generation function}\label{fig:generate_coeff_impl}
\end{figure}

Other convenience functionality also provided include plotting, function value perturbation, and function value evaluation based on domain span information that has been stored in the class instance. The function value evaluation is especially useful for experimenting and for the plotting functionality itself. This is because, the span was given to the constructor and does not need to be composed again with the inverse transform function.

\subsubsection{Model}

\noindent The SpectralSVR model is implemented in this module. There are two components in this module. First, there is the implementation of least-squares support vector regression (LSSVR) using Pytorch. This implementation is a reimplementation and modification of an LSSVR implementation using scikit-learn by \textcite{florencioLssvr2020}. The modifications are mainly concerned with using Pytorch in place of scikit-learn. Other modifications are concerned with the performance side of things, which are computing the kernel matrix in batches to avoid memory spikes and the ability to compute on hardware accelerators supported by Pytorch such as GPUs. The LSSVR implementation starts with the constructor which accepts model hyperparameters which include the regularization parameter, kernel function and any kernel parameters, verbosity level, kernel matrix computation batch size, and device to perform computations on. The recieved parameters are then stored as properties of the class instance. For the kernel specifically, it will be initialized with the correct default parameters when the fitting function is called. This is because if the kernel of choice is the radial basis function, the default scaling parameter needs to be computed from the training features as the square root of the sum of feature variance values.

The optimization function that fits the model to the training data solves \lccref{eq:lssvr_solution}. This function is shown in \lccref{fig:lssvr_fit_function_impl}. The first half of the function sets up the left and right hand side matrices. The left hand side matrix is constructed in part with the kernel matrix. The kernel function used is batched to reduce the peak memory footprint. These matrices are then used in the second half to compute the solution using the least squares solver function from Pytorch called lstq. Finally, the result is split into the biases and Lagrange multipliers which are then returned as a tuple. The optimizing function is wrapped in a more user-friendly function which allows the use of numpy arrays in addition to the Pytorch Tensor used in the optimizing function. The wrapper function is named fit. It conforms the inputs provided by the user into the expected types and matrix shapes of rows for samples and columns for input features or output targets. The processed inputs and outputs are also stored as support vectors for use during prediction. The fit function returns the class instance once training is done. This is for method chaining purposes for ease of use when using the fit function in a chain with other functions we will discuss next, such as the prediction function.

\begin{figure}[H]
  \centering
  \begin{lstlisting}[language=Python]
...
    def _optimize_parameters(
      self,
      X: torch.Tensor,
      y_values: torch.Tensor,
    ):

        A = torch.empty(
          (X.shape[0] + 1,) * 2,
          device=self.device,
          dtype=self.dtype,
        )
        A[1:, 1:] = self._batched_K(X, X)
        A[1:, 1:].diagonal().copy_(
            A[1:, 1:].diagonal()
            + torch.ones(
                (A[1:, 1:].shape[0],),
                device=self.device,
                dtype=self.dtype,
            ).to()
            / self.C
        )
        A[0, 0] = 0
        A[0, 1:] = 1
        A[1:, 0] = 1
        shape = np.array(y_values.shape)
        shape[0] += 1
        B = torch.empty(
          list(shape), device=self.device, dtype=self.dtype
        )
        B[0] = 0
        B[1:] = y_values

        solution: torch.Tensor = torch.linalg.lstsq(
            A.to(dtype=torch.float), B.to(dtype=torch.float)
        ).solution.to(dtype=self.dtype)

        b = solution[0, :]
        alpha = solution[1:, :]

        return (b, alpha)
...
  \end{lstlisting}
  \caption{Implementation of Least-squares Support Vector Regression fitting function}\label{fig:lssvr_fit_function_impl}
\end{figure}

The learned multipliers and bias parameters are then used for predicting outputs from unseen sample features. \Cref{eq:lssvr_prediction_vectorized} shows the computation that needs to be done to compute the predicted output. The prediction process first constructs both matrices on the left-hand side of the equation. This is done by computing the kernel matrix between the prediction features and the training features, otherwise known as support vectors. The prediction function therefore is very straight forward with an input of unseen features and the learned parameters. The function output is simply the predicted outputs. The implementation of this function is shown in \lccref{fig:lssvr_predict_function_impl}.
